You are the way you structurally talk: structural-temporal neighbourhoods of posts to characterize users in online forums

### 2016-04-28 doc/IEEEtran/paper_IEEEformat.tex
backup

\begin{definition}
Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $r$ from post $i$.
\end{definition}
% pros and cons
In the context of discussion threads, this definition has two limitations. First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written. In conversations, time plays an important role, therefore this is an important limitation. Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent.

@@

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{order_neighbourhood}
\includegraphics[width=0.45\textwidth]{breakpoints}
\caption{Illustration of order-based (left) and time-based (right) neighbourhoods. Grey nodes represent the ego post. Nodes with circles are those in the neighbourhood. Numbers indicate the order of the post. The order-based neighbourhood has parameters $r=3$, $n=6$. In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
\label{fig:cutpoints}
\end{figure*}
\subsection{Order-based}
Our first definition is based on the order in which posts are attached to the tree. 
\begin{definition}
Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.  
\end{definition}
An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints} (left).
This definition has two advantages over the \textit{structural neighbourhood}. First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are not only near to $i$ in the structure but also in time. Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite. The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to choose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.

@@

\subsection{Pruning}
At this point, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children. The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences. In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.

Figure~\ref{fig:colors} shows some real examples, from our dataset, of time-based neighbourhoods after colouring and pruning.

@@

We detect the order-based and time-based neighbourhood of every post as follows. A counter for every neighbourhood structure is created the first time it us detected. For every post, we extract its neighbourhood (order-based or time-based) colour and prune it as explained above (Section~\ref{sec:colouring_pruning}). Then we check whether it is isomorphic to some of the already seen neighbourhoods. If this is the case, then we increment the counter for the neighbourhood previously seen. Otherwise we create a new entry for the current neighbourhood. The computational cost of this operation has a upper bound of $\mathcal O(n^2)$, $n$ being the number of posts, for the case where each neighbourhood occurs only once. \footnote{TODO: It's actually much less, probably logarithmic, since the size dictionary grows incrementally}

\subsection{Agreement between neighbourhood methods}
The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures. Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times. Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to cover 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage. Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type. The reason why the frequency of the order-based neighbourhoods decay faster might be that the most frequent neighbourhood is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.

The divergent number of neighbourhoods between order-based and times-based brings us to the following question:  how differ the order-based and time-based neighbourhoods given to a same post? We re-labelled the order-based neighbourhoods and assigned them the label of their isomorphic time-based counterparts or the first unused number in the time-based index. This left us a set of 2183 trees. Figure~\ref{fig:confusion} shows, for every post, its time-based neighbourhood and its order-based neighbourhood. Those in the diagonal (34\%) see no difference between the time-based and order-based. Out of the diagonal, there are two rows indicating a high number of posts being assign the same order-based neighbourhood but very different time-based neighbourhoods. These order-based neighbourhoods are the number 1 and the number 50. These collapse of many time-based neighbourhoods into one is significant because it happens a lot among very frequent time-based neighbourhood. In the right plot of Figure ~\ref{fig:confusion} we can see other order-based neighbourhoods that collapse many of the most frequent time-based neighbourhoods, such as the 22, 17, 26, 29, 49, 56, 47 and 101.

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{confusion_time_order_gameofthrones}%
\includegraphics[width=0.5\textwidth]{confusion_time_order_gameofthrones_zoom}
\caption{Neighbourhood of posts from the point of view of order-based and time-based. Neighbourhoods with the same label in the order-based and the time-based axes are isomorphic. The points in the diagonal correspond to posts to whom the two methods assign exactly the same neighbourhood. Horizontal rows of points correspond to posts that are assigned many different time-based neighbourhoods but the same single order-based neighbourhood.}
\label{fig:confusion}
\end{figure*}

\subsection{Neighbourhood census}
Given a set of local structures and a graph under study, a census is a counting of how many times does each structure appear in the graph. In the context of online discussion forums, triad census has been sometimes used in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}. However, triads are poor objects when studying trees since there are only three possible structures with, at most, three nodes: a single vertex, a directed dyad (reply), star (two replies to the same post) and a chain (reply and reply to the reply). On the contrary, our neighbourhoods are potentially richer structures that capture the real structure of conversations.

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{order_vs_time_census_distribution}%
\includegraphics[width=0.5\textwidth]{order_vs_time_census_distribution_cum}
\caption{Comparison of census distributions with order-based and time-based neighbourhoods.}
\label{fig:census_distributions}
\end{figure*}

In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums. Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones. These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25). The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ... (27)

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{census_size_vs_freq_gameofthrones}
\caption{Size of neighbourhoods and their frequencies.}
\label{fig:census_size_vs_freq}
\end{figure}

This results lead us to think that the time-based neighbourhood is more convenient than the order-based. The main reason, we think, is that order seems to be a too dogmatic way to decide where a neighbourhood ends.

\subsection{Conversation-based clustering of users}
In this section, we cluster users based on the neighbourhoods of their posts. Our intuition is that some users like participating in some kind of discussion rather than other. Certainly, most part of the information necessary to understand the nature of a discussion in in its textual content. However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humour, irony, or context, we turn our attention towards the structure of the discussions, which can also contain some information. We work under the  hypothesis that the structural-neighbourhoods in which a user post in embedded reflect the kind of conversation in that part of the thread.

We limit our study to a set of 100 \textit{active users} who wrote more than 100 posts. For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the census, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$. We drop those feature columns that are zero for every user (these are neighbourhoods seen only around posts of users with low-level activity). To make the feature vector of a user independent on the number of posts, we transform the counts into percentages. And since some features have much higher percentages than others for most users, we scale and normalize the matrix so that every feature has mean 0 and variance 1. To avoid non-significant scores, we remove also the feature column corresponding to neighbourhoods that have a frequency less than 50 among the active users.

We use k-means to find the clusters, though one can use any other clustering method. Since the plot over the Within-Cluster Sum of Squares for $k=1,...20$ did not show any clear elbow we chose $k=3$ clusters so that clusters are easier to interpret. We did the clustering over a feature matrix with order-based neighbourhoods and another feature matrix with time-based neighbourhoods. Figure~\ref{fig:PCA} and shows the PCA projections of the users coloured by their assigned cluster.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{whiskers}
	\caption{Whiskers plots in the most relevant features (time-based)}
	\label{fig:PCA}
\end{figure}

@@

@@

\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\newpage
\section*{Appendix}
\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$ (Podemos)}
	\label{fig:census_orderbased}
\end{figure*}
\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_timebased_1}
	\includegraphics[width=0.8\textwidth]{census_timebased_2}
	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
	\caption{Most frequent time-based neighbourhoods (Podemos)}
	\label{fig:neighbourhoods_time}
\end{figure*}

>>>>

\begin{definition}
Given a tree graph $G$, the \textit{structure-based neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $r$ from post $i$.
\end{definition}
% pros and cons
In the context of discussion threads, this definition has two limitations. First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written. In conversations, time plays an important role, therefore this is an important limitation. Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent.

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{order_neighbourhood}
\includegraphics[width=0.45\textwidth]{breakpoints}
\caption{Illustration of order-based (left) and time-based (right) neighbourhoods. Grey nodes represent the ego post. Nodes with circles are those in the neighbourhood. Numbers indicate the order of the post. The order-based neighbourhood has parameters $r=3$, $n=6$. In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
\label{fig:cutpoints}
\end{figure}
\subsection{Order-based}
Our first definition is based on the order in which posts are attached to the tree. 
\begin{definition}
Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.  
\end{definition}
An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints} (left).
This definition has two advantages over the \textit{structural neighbourhood}. First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are not only near to $i$ in the structure but also in time. Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite. The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to choose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.

@@

\subsection{Pruning}
At this point, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children. The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences. In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.    
Figure~\ref{fig:colors} shows some real examples, from our dataset, of time-based neighbourhoods after colouring and pruning.

@@

We detect the order-based and time-based neighbourhood of every post as follows. A counter for every neighbourhood structure is created the first time it us detected. For every post, we extract its neighbourhood (order-based or time-based) colour and prune it as explained above (Section~\ref{sec:colouring_pruning}). Then we check whether it is isomorphic to some of the already seen neighbourhoods. If this is the case, then we increment the counter for the neighbourhood previously seen. Otherwise we create a new entry for the current neighbourhood. The computational cost of this operation has a upper bound of $\mathcal O(n^2)$, $n$ being the number of posts, for the case where each neighbourhood occurs only once. \footnote{TODO: It's actually much less, probably logarithmic, since the size dictionary grows incrementally}
Some of the anlysis in the next section requires that, if two type of neighbourhoods are isomorphics, they have the same label. Thus, we merge the dictionaries to obtain a global dictionary of neighbours. Figure XXX shows some frequent neighbourhoods that will be discussed later.

\subsection{Comparing neighbourhood methods}
In this section we compare different aspects of our three classes of neighbourhood, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, (c) the neighbourhood census obtained for each class in a same forum.

\subsubsection{Size and frequency distribution}
The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures. Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times. Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to cover 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage. Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type. The main reason why the frequency of the order-based neighbourhoods decay faster is that the most frequent neighbourhood (id XX) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.

\subsubsection{Discrepancies}
The divergent number of neighbourhoods between order-based and times-based suggests that posts with different time-based neighbourhoods have similar order-based neighbourhoods. Figure~\ref{fig:confusion} shows, for every post, its time-based neighbourhood and its order-based neighbourhood. Those in the diagonal (34\%) see no difference between the time-based and order-based. Out of the diagonal, there are two rows indicating a high number of posts being assign the same order-based neighbourhood but very different time-based neighbourhoods. These order-based neighbourhoods are the number 1 and the number 50. These collapse of many time-based neighbourhoods into one is significant because it happens a lot among very frequent time-based neighbourhood. Other order-based neighbourhoods that collapse many of the most frequent time-based neighbourhoods are the 22, 17, 26, 29, 49, 56, 47 and 101. 
The right figure \\TODO: expliquer... c'est time vs structure

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{confusion_time_order}%
\includegraphics[width=0.5\textwidth]{confusion_struct_time}
\includegraphics[width=0.5\textwidth]{confusion_time_order_zoom}%
\includegraphics[width=0.5\textwidth]{confusion_struct_time_zoom}
\caption{Neighbourhood of posts from the point of view of order-based and time-based. Neighbourhoods with the same label in the order-based and the time-based axes are isomorphic. The points in the diagonal correspond to posts to whom the two methods assign exactly the same neighbourhood. Horizontal rows of points correspond to posts that are assigned many different time-based neighbourhoods but the same single order-based neighbourhood.}
\label{fig:confusion}
\end{figure*}

\subsubsection{Neighbourhood census}
Given a set of local structures and a graph under study, a census is a counting of how many times does each structure appear in the graph. In the context of online discussion forums, triad census has been sometimes used in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}. However, triads are poor objects when studying trees since there are only three possible structures with, at most, three nodes: a single vertex, a directed dyad (reply), star (two replies to the same post) and a chain (reply and reply to the reply). On the contrary, our neighbourhoods are potentially richer structures that capture the real structure of conversations.

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones}%
\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones_cum}
\caption{Comparison of census distributions with order-based and time-based neighbourhoods.}
\label{fig:census_distributions}
\end{figure*}

In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums for the 20 most frequent neighbourhood of each type. Because some of the most frequent neighbourhoods in one method are also among the most frequent in the other methods, the total set contains 30 different neighbourhoods.

Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones. These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25). The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ... (27), suggesting that this method fails to capture some relevant structures.

@@

We conclude that, even if at the cost of a sparser set of structures, the time-based neighbourhood does a better job in capturing structures that have both a real meaning and a level of frequency that makes them relevant in the analysis.

\subsection{Conversation-based clustering of users}
In this section, we cluster users in the Podemos dataset based on the neighbourhoods of their posts. Our intuition is that some users like participating in some kind of discussion rather than other. Certainly, most part of the information necessary to understand the nature of a discussion in in its textual content. However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humour, irony, or context, we turn our attention towards the structure of the discussions, which can also contain some information. We work under the  hypothesis that the structural-neighbourhoods in which a user post in embedded reflect the kind of conversation in that part of the thread.

Our dataset contains a set of 100 \textit{active users} who wrote more than 100 posts. For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the census, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$. We drop those feature columns that are zero for every user (these are neighbourhoods seen only around posts of users with low-level activity). To make the feature vector of a user independent on the number of posts, we transform the counts into percentages. And since some features have much higher percentages than others for most users, we scale and normalize the matrix so that every feature has mean 0 and variance 1. To avoid non-significant scores, we remove also the feature column corresponding to neighbourhoods that have a frequency less than 50 among the active users.

We use k-means to find the clusters, though one can use any other clustering method. Since the plot over the Within-Cluster Sum of Squares for $k=1,...20$ did not show any clear elbow we chose $k=3$ clusters so that clusters are easier to interpret. We did the clustering over a feature matrix with order-based neighbourhoods and another feature matrix with time-based neighbourhoods.

Since the set of feature dimension is relatively large, we show in Figure \ref{fig:whiskers} how the users in a cluster are distributed for each feature. The features shown are the ones that have an bigger absolute value (in mean) for each cluster.

% Explicacion
\begin{itemize}
\item The green cluster (attraction for X, X, repulsion XXX) correspond to users that...
\item The red cluster...
\item The blue cluster...
\end{itemize}

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{whiskers}
	\caption{Whiskers plots in the most relevant features (time-based)}
	\label{fig:whiskers}
\end{figure}

@@

\section{Future work}
In this paper we focused on the comparative analysis between three kind of neighbourhoods in order to demonstrate that the time-based neighbourhoods are able to capture relevant structures that other definitions of neighbourhood cannot. We also showed that it can be used to characterise users according to the kind of conversation in which they participate.

We can see many applications of this kind of neighbourhood.

\begin{itemize}
\item  Caracterise reactions to user posts: instead of capturing the neighbours all around a post, we can limit we neighbourhood to the descendants. This would allow to cluster users from a slightly (but maybe important) point of view. 
\item The boundaries of our time-based neighbourhood are decided by the changepoint algorithm PELT. How sensible is time-based neighbourhood to the choice of the algorithm?  
\item Merge more neighbourhoods with sociological criteria.
\item Use different coulours to represent different phenomena.
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{dictionary1}
	\includegraphics[width=0.8\textwidth]{dictionary2}
	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
	\caption{Neighbourhoods reference list}
	\label{fig:census_orderbased}
\end{figure*}

@@

\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\newpage
\section*{Appendix 2}

\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$ (Podemos)}
	\label{fig:census_orderbased}
\end{figure*}
\begin{figure*}
	\centering
	\includegraphics[width=0.8\textwidth]{census_timebased_1}
	\includegraphics[width=0.8\textwidth]{census_timebased_2}
	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
	\caption{Most frequent time-based neighbourhoods (Podemos)}
	\label{fig:neighbourhoods_time}
\end{figure*}
